Parameter estimation

### 2023-06-21 SFE-article/Article_Model_Parameters.tex
20.06
21.06

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
The aim of this study was to investigate the supercritical extraction process of the caraway oil from caraway seeds from a partially filled extractor with a fixed bed that was operated under multiple constant operating conditions. The dataset collected from the experiments consists of four experiments conducted at different operating conditions: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. To describe the fluid-solid extraction process with $CO_2$ as a solvent, a distributed-parameter model is used. It is assumed that the flow is uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. To model this two-dimensional case, the concept of quasi-one-dimensional flow is applied. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, saturation concentration, and an initial state estimate. These parameters are estimated from the Peng-Robinson equation of state and by applying the maximum likelihood estimation method on yield data under the normal error assumption.

@@

\subsection{Governing equations} \label{CH:Governing_equations_chapter}
The governing equation for a quasi-one-dimensional compressible flow in Cartesian coordinates can be found in the Appendix \ref{CH: Gouverning equations} and in the work of \citet{Anderson1995}.  Quasi-one-dimensional flow is a fluid flow characterized by the assumption that the flow properties remain uniform across any given cross-section of the flow. This assumption is made when there is a variation in the cross-sectional area of the flow channel, such as an irregular shape or partial filling of an extractor. In such cases, the flow is considered to be quasi-one-dimensional because the velocity and other flow properties are assumed to vary only in the direction of flow.

@@

{\footnotesize
	\begin{align}
		\label{EQ: CompressibleEuler_1}
		\cfrac{\partial \left( \rho_f A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v \right)}{\partial z} &= 0 \\
		\cfrac{\partial \left( \rho_f v A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v^2 \right)}{\partial z} &= -A_f \cfrac{\partial P}{\partial z} \label{EQ: CompressibleEuler_2} \\
		\cfrac{\partial \left( \rho_f e A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v e\right)}{\partial z} &= -P\cfrac{\left( A_f v \right)}{\partial z} + \cfrac{\partial}{\partial z} \left( \cfrac{\partial T}{\partial z} \right)   
		\label{EQ: CompressibleEuler_3}
	\end{align}  
}

where $\rho_f$ is the density of the fluid, $A_f$ is the function which describe change of the cross-section, $v$ is the velocity, $P$ is the total pressure, $e$ is the internal energy of the fluid, $t$ is time and $z$ is the spacial direction.

@@

\begin{figure}[!h]
	\centering
	\resizebox{0.95\columnwidth}{!}{%
		\begin{tikzpicture}[]
			\draw (0,2) -- (9,2);	% Top line
			\draw (0,0) -- (9,0); 	% Bottom line
			\draw (4.5,0) -- (4.5,2); 	% Bottom line
			\node at (4,1.5) {$\rho_f$};
			\node at (5.5,1.5) {$\rho_f+\delta\rho_f$};
			\node at (4,1.0) {$P$};
			\node at (5.5,1.0) {$P+\delta P$};
			\node at (4,0.5) {$T$};
			\node at (5.5,0.5) {$T+\delta T$};
			\node at (1,1.0) {$v \rightarrow$};
			\node at (8,1.0) {$v + \delta v \rightarrow$};
	\end{tikzpicture} }
	\caption{Small discontinuity in one-dimensional flow}
	\label{fig: Discontinuity_slow_flow}
\end{figure}

@@

{\footnotesize
	\begin{align*}
		&\rho_f \delta v + v \delta \rho_f + \delta \rho_f \delta v = 0 \\
		&\delta P = \delta v \delta \rho_f
	\end{align*}
}

@@

{\footnotesize
	\begin{equation*}
		\lim_{\rho_f v \rightarrow 0} \rho_f \delta v + v \delta \rho_f + \delta \rho_f \delta v = 0 / \delta \rho_f \rightarrow \cfrac{d v}{d \rho_f} = - \cfrac{v}{\rho_f}
	\end{equation*}
}

@@

{\footnotesize
	\begin{equation} \label{EQ: Pressure_Velocity}
		\cfrac{d v}{d \rho_f} = - \cfrac{d v}{dP} \cfrac{d P}{d \rho_f} = -\cfrac{1}{\rho v} \cfrac{dP}{d\rho_f} = -\cfrac{v}{\rho_f}
	\end{equation}
}

Suppose the flow is presumed to be isentropic, $dP/d\rho = c^2$, so $v^2=c^2$, where $c$ is the speed of sound. This can be interpreted as a small pressure wave propagating with the speed of sound relative to the flow. Moreover, if the flow velocity is relatively low, all pressure changes are hydrodynamic (due to velocity motion) rather than thermodynamic which leads to $\partial \rho / \partial P \approx 0$. In other words, the small changes in pressure due to flow velocity changes do not change the density. This has a secondary effect -- the speed of sound in the fluid is $\partial P/\partial \rho = \infty$ in this instance. So there is an infinite speed of sound, which makes the equations elliptic in nature. It can be deduced that at the isothermal conditions, the density in the system propagates with the same speed as pressure since they are both connected through the equation of state. The details of the Mach-number analysis are presented in Appendix \ref{CH:Low_Mach_chapter}.

@@

@@

\clearpage
\subsection{Solution of the parameter estimation starting from another initial guess} \label{CH: Profiles_2}
\subfile{Sections/Parameter_estimation_solution_from_another_IG}

>>>>

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
The aim of this study was to investigate the supercritical extraction process of the caraway oil from caraway seeds from a partially filled extractor with a fixed bed that was operated under multiple constant operating conditions. The dataset collected from the experiments consists of four experiments conducted at different operating conditions: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. To describe the fluid-solid extraction process with $CO_2$ as a solvent, a distributed-parameter model is used. To reduce the number of spatial dimensions, the concept of quasi-one-dimensional flow is applied. It is assumed that the flow is uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, saturation concentration, and an initial state estimate. These parameters are estimated from the Peng-Robinson equation of state and by applying the maximum likelihood estimation method on yield data under the normal error assumption.

@@

\subsection{Governing equations} \label{CH:Governing_equations_chapter}
The governing equation \footnote{For the sake of clarity of the process model, different colors have been used in the equations to indicate: 
	{\color{red}control variables},
	{\color{blue}state variables},
	{\color{orange}variables} and
	{\color{magenta}parameters}.} for a quasi-one-dimensional compressible flow in Cartesian coordinates can be found in the Appendix \ref{CH: Gouverning equations} and in the work of \citet{Anderson1995}.  Quasi-one-dimensional flow is a fluid flow characterized by the assumption that the flow properties remain uniform across any given cross-section of the flow. This assumption is made when there is a variation in the cross-sectional area of the flow channel, such as an irregular shape or partial filling of an extractor. In such cases, the flow is considered to be quasi-one-dimensional because the velocity and other flow properties are assumed to vary only in the direction of flow.

@@

{\footnotesize
	\begin{align}
		\label{EQ: CompressibleEuler_1}
		\cfrac{\partial \left( {\color{orange}\rho_f} {\color{magenta}A_f}(z) \right) }{\partial t} + \cfrac{\partial \left( {\color{orange}\rho_f} {\color{magenta}A_f}(z) v \right)}{\partial z} &= 0 \\
		\cfrac{\partial \left( {\color{orange}\rho_f} v {\color{magenta}A_f}(z) \right) }{\partial t} + \cfrac{\partial \left( {\color{orange}\rho_f} {\color{magenta}A_f}(z) v^2 \right)}{\partial z} &= -{\color{magenta}A_f}(z) \cfrac{\partial {\color{blue}P}}{\partial z} \label{EQ: CompressibleEuler_2} \\
		\cfrac{\partial \left( {\color{orange}\rho_f} {\color{blue}e} {\color{magenta}A_f}(z) \right) }{\partial t} + \cfrac{\partial \left( {\color{orange}\rho_f} {\color{magenta}A_f}(z) v {\color{blue}e}\right)}{\partial z} &= -{\color{blue}P}\cfrac{\left( {\color{magenta}A_f}(z) v \right)}{\partial z} + \cfrac{\partial}{\partial z} \left( \cfrac{\partial {\color{blue}T}}{\partial z} \right)   
		\label{EQ: CompressibleEuler_3}
	\end{align}  
}

where ${\color{orange}\rho_f}$ is the density of the fluid, ${\color{magenta}A_f}(z)$ is the function which describe change of the cross-section, $v$ is the velocity, ${\color{blue}P}$ is the total pressure, $e$ is the internal energy of the fluid, $t$ is time and $z$ is the spacial direction.

@@

\begin{figure}[!h]
	\centering
	\resizebox{0.95\columnwidth}{!}{%
		\begin{tikzpicture}[]
			\draw (0,2) -- (9,2);	% Top line
			\draw (0,0) -- (9,0); 	% Bottom line
			\draw (4.5,0) -- (4.5,2); 	% Bottom line
			\node at (4,1.5) {${\color{orange}\rho_f}$};
			\node at (5.5,1.5) {${\color{orange}\rho_f}+\delta{\color{orange}\rho_f}$};
			\node at (4,1.0) {$P$};
			\node at (5.5,1.0) {${\color{blue}P}+\delta {\color{blue}P}$};
			\node at (4,0.5) {$T$};
			\node at (5.5,0.5) {${\color{blue}T}+\delta {\color{blue}T}$};
			\node at (1,1.0) {$v \rightarrow$};
			\node at (8,1.0) {$v + \delta v \rightarrow$};
	\end{tikzpicture} }
	\caption{Small discontinuity in one-dimensional flow}
	\label{fig: Discontinuity_slow_flow}
\end{figure}

@@

{\footnotesize
	\begin{align*}
		&{\color{orange}\rho_f} \delta v + v \delta {\color{orange}\rho_f} + \delta {\color{orange}\rho_f} \delta v = 0 \\
		&\delta {\color{blue}P} = \delta v \delta {\color{orange}\rho_f}
	\end{align*}
}

@@

{\footnotesize
	\begin{equation*}
		\lim_{{\color{orange}\rho_f} v \rightarrow 0} {\color{orange}\rho_f} \delta v + v \delta {\color{orange}\rho_f} + \delta {\color{orange}\rho_f} \delta v = 0 / \delta {\color{orange}\rho_f} \rightarrow \cfrac{d v}{d {\color{orange}\rho_f}} = - \cfrac{v}{{\color{orange}\rho_f}}
	\end{equation*}
}

@@

{\footnotesize
	\begin{equation} \label{EQ: Pressure_Velocity}
		\cfrac{d v}{d {\color{orange}\rho_f}} = - \cfrac{d v}{d{\color{blue}P}} \cfrac{d P}{d {\color{orange}\rho_f}} = -\cfrac{1}{\rho v} \cfrac{d{\color{blue}P}}{d{\color{orange}\rho_f}} = -\cfrac{v}{{\color{orange}\rho_f}}
	\end{equation}
}

Suppose the flow is presumed to be isentropic, $d{\color{blue}P}/d{\color{orange}\rho_f} = c^2$, so $v^2=c^2$, where $c$ is the speed of sound. This can be interpreted as a small pressure wave propagating with the speed of sound relative to the flow. Moreover, if the flow velocity is relatively low, all pressure changes are hydrodynamic (due to velocity motion) rather than thermodynamic which leads to $\partial {\color{orange}\rho_f} / \partial {\color{blue}P} \approx 0$. In other words, the small changes in pressure due to flow velocity changes do not change the density. This has a secondary effect -- the speed of sound in the fluid is $\partial {\color{blue}P}/\partial {\color{orange}\rho_f} = \infty$ in this instance. So there is an infinite speed of sound, which makes the equations elliptic in nature. It can be deduced that at the isothermal conditions, the density in the system propagates with the same speed as pressure since they are both connected through the equation of state. The details of the Mach-number analysis are presented in Appendix \ref{CH:Low_Mach_chapter}.

@@

\subsection{Maximum Likelihood} \label{CH: ML}
\subfile{Sections/Likelihood}

@@

%\subsection{Solution of the parameter estimation starting from another initial guess} \label{CH: Profiles_2}
%\subfile{Sections/Parameter_estimation_solution_from_another_IG}
